---
title: "COVID 19 CABA // ANALYSIS NEIGHBORHOODS"
output: html_notebook
---

### Load libraries

```{r}
library(tidyverse) #for all functions in tidyverse
library(lubridate) #for dates format

```

### Import datasets

```{r}
#Confirmed Covid cases
covid_CABA_confirmed <- read.csv("clean_covid_CABA_confirmed.csv", stringsAsFactors = FALSE)

#Deaths by Covid
covid_CABA_dead <- read.csv("clean_covid_CABA_dead.csv", stringsAsFactors = FALSE)

#Population data
population_data <- read.csv("clean_population_data.csv", stringsAsFactors = FALSE)
population_data$X <- NULL

#Population by neighborhood according to 2010 census
population_by_neighborhood <- read.csv("caba_pob_barrios_2010.csv", stringsAsFactors = FALSE)

```

## ANALYSIS

### NEIGHBORHOODS

Create a dataframe with the number of cases per neighborhood and aggregate the population per neighborhood:

```{r}

#Crear dataframes temnporales para casos confirmados y muertes y renombrar columnas

cases_by_neighborhood_temp <- data.frame(table(covid_CABA_confirmed$neighborhood))
deaths_by_neighborhood_temp <- data.frame(table(covid_CABA_dead$neighborhood))

colnames(cases_by_neighborhood_temp) <-c("neighborhood","confirmed_cases")
colnames(deaths_by_neighborhood_temp) <-c("neighborhood","deaths")

#Create a dataframe with:
#
#1. Cases
#2. Cases per 1M
#3. % of population infected
#4. Deaths
#5. Deaths per 1M

cases_by_neighborhood <- population_by_neighborhood %>%
  left_join(cases_by_neighborhood_temp, by = "neighborhood")

cases_by_neighborhood <- cases_by_neighborhood %>%
  mutate(confirmed_by_1M = round(1000000 * confirmed_cases / population))

cases_by_neighborhood <- cases_by_neighborhood %>%
  mutate(percentage_confirmed = round(confirmed_by_1M / 10000,2))

cases_by_neighborhood <- cases_by_neighborhood %>%
  left_join(deaths_by_neighborhood_temp, by = "neighborhood")

cases_by_neighborhood <- cases_by_neighborhood %>%
  mutate(deaths_by_1M = round(1000000 * deaths / population))

#Sort

cases_by_neighborhood <- arrange(cases_by_neighborhood, neighborhood)

#Check dataframe

cases_by_neighborhood


```



Save confirmed cases and deaths occurring in unknown neighborhoods:


```{r}
confirmed_unknown_neighborhood = cases_by_neighborhood_temp[1,2]
deaths_unknown_neighborhood = deaths_by_neighborhood_temp[1,2]

```

### Calculate cumulative cases and cumulative cases per million, per month, per neighborhood
```{r}

#Group by year and month

cumulative_cases_neighborhood <- covid_CABA_confirmed %>%
  group_by(neighborhood, 'year' = year(case_date), 'month' = format.Date(case_date, "%m")) %>%
  tally()


#Remove cases without identified neighborhood

cumulative_cases_neighborhood <- cumulative_cases_neighborhood[!(is.na(cumulative_cases_neighborhood$neighborhood) | cumulative_cases_neighborhood$neighborhood==""), ]

#Create a column for months and delete unnecessary year column

cumulative_cases_neighborhood$month <- as.character(paste(cumulative_cases_neighborhood$year, cumulative_cases_neighborhood$month))

cumulative_cases_neighborhood$year <- NULL

#Create the table with pivot_wider, taking the months as column name and set the months with no cases to 0

cumulative_cases_neighborhood <- cumulative_cases_neighborhood %>% pivot_wider(names_from = month, values_from = n)

cumulative_cases_neighborhood[is.na(cumulative_cases_neighborhood)] <- 0

#Transpose, calculate the cumulative sum and transpose again. Then re-add the column with the names of the neighborhoods

cumulative_cases_neighborhood <- as.data.frame(t(apply(t(cumulative_cases_neighborhood[,-1]),2, cumsum))) %>% add_column("neighborhood"=population_by_neighborhood$neighborhood, .before = '2020 03')

#Put the names of the months in columns

names(cumulative_cases_neighborhood)[-1] <- paste(month.abb[as.numeric(sub(".* ", "", names(cumulative_cases_neighborhood[-1])))], substr(names(cumulative_cases_neighborhood[-1]), 1,4), sep = " ")


#Calculate cumulative cases per 1M

cumulative_cases_1M_neighborhood <- round(1000000 * cumulative_cases_neighborhood[-1] / t(population_by_neighborhood$population)) %>%
  add_column("neighborhood"=population_by_neighborhood$neighborhood, .before = 'Mar 2020')

#Check dataframes

cumulative_cases_neighborhood
cumulative_cases_1M_neighborhood

#Calculate number and % of cases without commune identified until Feb 2021 (to specify in map)

nrow(filter(covid_CABA_confirmed, is.na(commune), case_date <= "2021-02-28"))

nrow(filter(covid_CABA_confirmed, is.na(commune), case_date <= "2021-02-28")) * 100 /
nrow(filter(covid_CABA_confirmed, case_date <= "2021-02-28"))


```

### Repeat the procedure for cumulative deaths/cumulative deaths per 1M, per month, per neighborhood


```{r}

#Group by year and month

cumulative_deaths_neighborhood <- covid_CABA_dead %>%
  group_by(neighborhood, 'year' = year(case_date), 'month' = format.Date(case_date, "%m")) %>%
  tally()

#Remove cases without identified neighborhood

cumulative_deaths_neighborhood <- cumulative_deaths_neighborhood[!(is.na(cumulative_deaths_neighborhood$neighborhood) | cumulative_deaths_neighborhood$neighborhood==""), ] 


#Create a column for months and delete unnecessary year column

cumulative_deaths_neighborhood$month <- as.character(paste(cumulative_deaths_neighborhood$year, cumulative_deaths_neighborhood$month))

cumulative_deaths_neighborhood$year <- NULL

#Create the table with pivot_wider, taking the months as column name and set the months with no cases to 0

cumulative_deaths_neighborhood <- cumulative_deaths_neighborhood %>% pivot_wider(names_from = month, values_from = n)

cumulative_deaths_neighborhood[is.na(cumulative_deaths_neighborhood)] <- 0

# Remove neighborhood column and sort by month using column names

#// The dataframe columns are unsorted because the first neighborhood in alphabetical name -used before to pivot- didn't register any deaths for the first two months) //#

cumulative_deaths_neighborhood <-  cumulative_deaths_neighborhood[-1]

cumulative_deaths_neighborhood <-  cumulative_deaths_neighborhood[, order(names(cumulative_deaths_neighborhood))]

#Transpose, calculate the cumulative sum and transpose again. Then re-add the column with the names of the neighborhoods

cumulative_deaths_neighborhood <- as.data.frame(t(apply(t(cumulative_deaths_neighborhood),2, cumsum))) %>% add_column("neighborhood"=population_by_neighborhood$neighborhood, .before = '2020 03')
  

#Put the names of the months in columns

names(cumulative_deaths_neighborhood)[-1] <- paste(month.abb[as.numeric(sub(".* ", "", names(cumulative_deaths_neighborhood[-1])))], substr(names(cumulative_deaths_neighborhood[-1]), 1,4), sep = " ")


#Calculate cumulative deaths per 1M

cumulative_deaths_1M_neighborhood <- round(1000000 * cumulative_deaths_neighborhood[-1] / t(population_by_neighborhood$population)) %>% add_column("neighborhood"=population_by_neighborhood$neighborhood, .before = 'Mar 2020')

#Check dataframes

cumulative_deaths_neighborhood
cumulative_deaths_1M_neighborhood


```

### Export tables

```{r}


#Cases by neighborhood
write.csv(cases_by_neighborhood, "final_cases_by_neighborhood.csv")

#Cumulative cases per neighborhood
write.csv(cumulative_cases_neighborhood, "final_cumulative_cases_neighborhood.csv")

#Cumulative cases by 1M per neighborhood
write.csv(cumulative_cases_1M_neighborhood, "final_cumulative_cases_1M_neighborhood.csv")

#Cumulative deaths per neighborhood
write.csv(cumulative_deaths_neighborhood, "final_cumulative_deaths_neighborhood.csv")

#Cumulative deaths by 1M per neighborhood
write.csv(cumulative_deaths_1M_neighborhood, "final_cumulative_deaths_1M_neighborhood.csv")


 no_commune_until_feb_2021 <- filter(covid_CABA_confirmed, is.na(commune), case_date <= "2021-02-28")



```
```




