---
title: "COVID 19 CABA // ANALYSIS COMMUNES"
output: html_notebook
---

### Load libraries

```{r}
library(tidyverse) #for all functions in tidyverse
library(lubridate) #for dates format
#library(zoo) #for moving averages
#library('skimr')

```

### General functions

```{r}

#Percentage
perc <- function (quant,total) {
  p <- quant/total
  p - 1
  p * 100
}
```

### Import datasets

```{r}
#Confirmed Covid cases
covid_CABA_confirmed <- read.csv("clean_covid_CABA_confirmed.csv", stringsAsFactors = FALSE)

#Deaths by Covid
covid_CABA_dead <- read.csv("clean_covid_CABA_dead.csv", stringsAsFactors = FALSE)

#Population data

population_data <- read.csv("clean_population_data.csv", stringsAsFactors = FALSE)
population_data$X <- NULL

```

## ANALYSIS

### COMMUNES

Create a dataframe with the number of cases per commune and add the population per commune:

```{r}

#Create temporary dataframes for confirmed cases and deaths and rename columns

cases_by_commune_temp <- data.frame(table(covid_CABA_confirmed$commune))
deaths_by_commune_temp <- data.frame(table(covid_CABA_dead$commune))

colnames(cases_by_commune_temp) <-c("commune","confirmed_cases")
cases_by_commune_temp$commune <- as.numeric(cases_by_commune_temp$commune)

colnames(deaths_by_commune_temp) <-c("commune","deaths")
deaths_by_commune_temp$commune <- as.numeric(deaths_by_commune_temp$commune)

#Add to dataframe
#
#1. Cases
#2. Cases by 1M
#3. Deaths
#4. Deaths by 1M
#5. Mortality rate

cases_by_commune <- population_data %>%
  left_join(cases_by_commune_temp, by = "commune")

cases_by_commune <- cases_by_commune %>%
  mutate(confirmed_by_1M = round(1000000 * confirmed_cases / population)) %>%
  arrange(confirmed_by_1M)

cases_by_commune <- cases_by_commune %>%
  left_join(deaths_by_commune_temp, by = "commune")

cases_by_commune <- cases_by_commune %>%
  mutate(deaths_by_1M = round(1000000 * deaths / population))

cases_by_commune <- cases_by_commune %>%
  mutate(mortality_rate = round(deaths / confirmed_cases * 100))

#Sort

cases_by_commune <- arrange(cases_by_commune, commune)

#Check dataframe

cases_by_commune

```

Save confirmed cases and deaths occurring in unknown communes:

```{r}

confirmed_unknown_commune = nrow(covid_CABA_confirmed[covid_CABA_confirmed$commune == '',])

deaths_unknown_commune = nrow(covid_CABA_dead[covid_CABA_dead$commune == '',])

```


### Calculate cumulative cases and cumulative cases per million, per month, per commune

```{r}

#Group by year and month

cumulative_cases_commune <- covid_CABA_confirmed %>%
  group_by(commune, 'year' = year(case_date), 'month' = format.Date(case_date, "%m")) %>%
  tally()

#Remove cases without identified commune

cumulative_cases_commune <- cumulative_cases_commune[!(is.na(cumulative_cases_commune$commune) | cumulative_cases_commune$commune==""), ]

#Create a column for months and delete unnecessary year column

cumulative_cases_commune$month <- as.character(paste(cumulative_cases_commune$year, cumulative_cases_commune$month))

cumulative_cases_commune$year <- NULL


#Create the table with pivot_wider, taking the months as column name and set the months with no cases to 0

cumulative_cases_commune <- cumulative_cases_commune %>% pivot_wider(names_from = month, values_from = n)

cumulative_cases_commune[is.na(cumulative_cases_commune)] <- 0

#Transpose, calculate the cumulative sum and transpose again. Then re-add the column with the names of the communes

cumulative_cases_commune <- as.data.frame(t(apply(t(cumulative_cases_commune[,-1]),2, cumsum))) %>% add_column("commune"=population_data$commune, .before = '2020 03')

#Put the names of the months in columns

names(cumulative_cases_commune)[-1] <- paste(month.abb[as.numeric(sub(".* ", "", names(cumulative_cases_commune[-1])))], substr(names(cumulative_cases_commune[-1]), 1,4), sep = " ")


#Calculate cumulative cases per 1M

cumulative_cases_1M_commune <- round(1000000 * cumulative_cases_commune[-1] / t(population_data$population)) %>%
  add_column("commune"=population_data$commune, .before = 'Mar 2020')


#Check dataframes

cumulative_cases_commune
cumulative_cases_1M_commune


```


### Repeat the procedure for cumulative deaths/cumulative deaths per 1M, per month, per commune


```{r}

#Group by year and month

cumulative_deaths_commune <- covid_CABA_dead %>%
  group_by(commune, 'year' = year(case_date), 'month' = format.Date(case_date, "%m")) %>%
  tally()

#Remove cases without identified commune

cumulative_deaths_commune <- cumulative_deaths_commune[!(is.na(cumulative_deaths_commune$commune) | cumulative_deaths_commune$commune==""), ] 


#Create a column for months and delete unnecessary year column

cumulative_deaths_commune$month <- as.character(paste(cumulative_deaths_commune$year, cumulative_deaths_commune$month))

cumulative_deaths_commune$year <- NULL

#Create the table with pivot_wider, taking the months as column name and set the months with no cases to 0


cumulative_deaths_commune <- cumulative_deaths_commune %>% pivot_wider(names_from = month, values_from = n)

cumulative_deaths_commune[is.na(cumulative_deaths_commune)] <- 0

# Remove commune column and sort by month using column names

cumulative_deaths_commune <-  cumulative_deaths_commune[-1]

cumulative_deaths_commune <-  cumulative_deaths_commune[, order(names(cumulative_deaths_commune))]

#Transpose, calculate the cumulative sum and transpose again. Then re-add the column with the names of the communes

cumulative_deaths_commune <- as.data.frame(t(apply(t(cumulative_deaths_commune),2, cumsum))) %>% add_column("commune"=population_data$commune, .before = '2020 03')
  

#Put the names of the months in columns

names(cumulative_deaths_commune)[-1] <- paste(month.abb[as.numeric(sub(".* ", "", names(cumulative_deaths_commune[-1])))], substr(names(cumulative_deaths_commune[-1]), 1,4), sep = " ")


#Calculate cumulative deaths per 1M

cumulative_deaths_1M_commune <- round(1000000 * cumulative_deaths_commune[-1] / t(population_data$population)) %>% add_column("commune"=population_data$commune, .before = 'Mar 2020')

#Check dataframes

cumulative_deaths_commune
cumulative_deaths_1M_commune

```

## Is there any correlation between final cases/deaths figures and income, overcrowding or poverty?

```{r}


correlation_table <- cases_by_commune[, c(4,5,6,7,8,9,10,11)]
cor(correlation_table)


```

## There's no strong correlation with the final figures. But if we try month by month?


```{r}

#Create a dataframe to store monthly cumulative correlations between cases/deaths and poverty

correlation_poverty <- cumulative_cases_1M_commune[(0:4),]
correlation_poverty[,] <- 0

colnames(correlation_poverty)[1] <- ''
correlation_poverty[1,1] <- 'cases'
correlation_poverty[2,1] <- 'cases_1M'
correlation_poverty[3,1] <- 'deaths'
correlation_poverty[4,1] <- 'deaths_1M'

#Populate the dataframe by correlating monthly cumulative figures with poverty
#
#NOTE: probably there's a better way of doing this without using a "for" loop

for (i in 2:ncol(cumulative_cases_commune)) { #from mar 2020 to jan 2022
   
   correlation_poverty[1,i] <- cor(cumulative_cases_commune[i],population_data$poverty) #correlate with cases
  
  correlation_poverty[2,i] <- cor(cumulative_cases_1M_commune[i],population_data$poverty) #correlate with cases by 1M
  
  correlation_poverty[3,i] <- cor(cumulative_deaths_commune[i],population_data$poverty) #correlate with deaths
  
  correlation_poverty[4,i] <- cor(cumulative_deaths_1M_commune[i],population_data$poverty) #correlate with deaths by 1M
}

#Check correlation

correlation_poverty

```

## Repeat procedure to correlate between poverty and overcrowding

```{r}

#Create a dataframe to store monthly cumulative correlations between cases/deaths and overcrowding

correlation_overcrowding <- cumulative_cases_1M_commune[(0:4),]
correlation_overcrowding[,] <- 0

colnames(correlation_overcrowding)[1] <- ''
correlation_overcrowding[1,1] <- 'cases'
correlation_overcrowding[2,1] <- 'cases_1M'
correlation_overcrowding[3,1] <- 'deaths'
correlation_overcrowding[4,1] <- 'deaths_1M'

#Populate the dataframe by correlating monthly cumulative figures with overcrowding
#
#NOTE: probably there's a better way of doing this without using a "for" loop

for (i in 2:ncol(cumulative_cases_commune)) { #from mar 2020 to jan 2022
  correlation_overcrowding[1,i] <- cor(cumulative_cases_commune[i],population_data$overcrowding) #correlate with cases
  
  correlation_overcrowding[2,i] <- cor(cumulative_cases_1M_commune[i],population_data$overcrowding) #correlate with cases by 1M
  
  correlation_overcrowding[3,i] <- cor(cumulative_deaths_commune[i],population_data$overcrowding) #correlate with deaths
  
  correlation_overcrowding[4,i] <- cor(cumulative_deaths_1M_commune[i],population_data$overcrowding) #correlate with deaths by 1M
}

#Check correlation

correlation_overcrowding

```

### Export tables

```{r}

#Cases by commune
write.csv(cases_by_commune, "final_cases_by_commune.csv")

#Cumulative cases per commune
write.csv(cumulative_cases_commune,"final_cumulative_cases_by_commune.csv")

#Cumulative cases by 1M per commune
write.csv(cumulative_cases_1M_commune,"final_cumulative_cases_1M_by_commune.csv")

#Cumulative deaths per commune
write.csv(cumulative_deaths_commune,"final_cumulative_deaths_by_commune.csv")

#Cumulative deaths by 1M per commune
write.csv(cumulative_deaths_1M_commune,"final_cumulative_deaths_1M_by_commune.csv")

#Monthly correlation between cases/deaths and poverty
write.csv(correlation_poverty, "final_correlation_poverty.csv")

#Monthly correlation between cases/deaths and overcrowding
write.csv(correlation_overcrowding, "final_correlation_overcrowding.csv")


```