---
title: "COVID 19 CABA // GENERAL ANALYSIS CABA: CASES AND VACCINATION"
output: html_notebook
---

### Load libraries

```{r}
library(tidyverse) #for all functions in tidyverse
library(zoo) #for moving averages
library(lubridate) #for dates format

```

### General functions

```{r}

#Percentage
perc <- function (quant,total) {
  p <- quant/total
  p - 1
  p * 100
}

```

### Import datasets & create main variables

```{r}

#Confirmed Covid cases
covid_CABA_confirmed <- read.csv("clean_covid_CABA_confirmed.csv", stringsAsFactors = FALSE)

#Deaths by Covid
covid_CABA_dead <- read.csv("clean_covid_CABA_dead.csv", stringsAsFactors = FALSE)

#Population data
population_data <- read.csv("clean_population_data.csv", stringsAsFactors = FALSE)
population_data$X <- NULL

total_population <- sum(population_data$population)

#Vaccination
vaccination_all <- read.csv("clean_vaccination.csv", stringsAsFactors = FALSE)
vaccination_all$X <- NULL
vaccination_all[is.na(vaccination_all)] <- 0

```

## ANALYSIS

### TOTAL CASES

Create a dataframe of cases and deaths per day:

```{r}

#Create a dataframe with:
#
#1. Cases
#2. Cumulative cases
#3. Deaths per day
#4. Cumulative deaths
#5. Mortality rate
#6. Rolling average deaths for the last 7 days

cases_by_date <- data.frame(table(covid_CABA_confirmed$case_date))
deaths_by_date_temp <- data.frame(table(covid_CABA_dead$death_date))
  
colnames(cases_by_date) <-c("date","confirmed_cases")
colnames(deaths_by_date_temp) <-c("date","deaths")

cases_by_date <- cases_by_date %>%
  left_join(deaths_by_date_temp, by = "date")

#Add missing days
cases_by_date <- cases_by_date %>%
  mutate(date = as.Date(date)) %>%
  complete(date = seq.Date(min(date), max(date), by="day"))

#Add average number of cases and deaths per 7 days
cases_by_date[is.na(cases_by_date)] <- 0

cases_by_date <- cases_by_date %>%
  mutate(avg_cases_per_week = round(rollmeanr(confirmed_cases, k = 7, fill = NA),2),
avg_deaths_per_week = round(rollmeanr(deaths, k = 7, fill = NA),2))

#Add cumulative cases, cumulative deaths and mortality
cases_by_date <- cases_by_date %>%
  mutate(cumulative_cases = cumsum(confirmed_cases),
cumulative_deaths = cumsum(deaths))

cases_by_date$mortality <- round(cases_by_date$cumulative_deaths * 100 / cases_by_date$cumulative_cases,2)

```

### Create a function to calculate cases, deaths and mortality between specific dates

```{r}

cases_between_dates <- function (initial_date, final_date) {
  
  cumulative_cases_deaths_mortality <- c(0,0,0)
  
  date_range <- subset(cases_by_date, cases_by_date$date>=initial_date & cases_by_date$date<= final_date)
  
  cumulative_cases_deaths_mortality[1] <- (sum(date_range$confirmed_cases))
  cumulative_cases_deaths_mortality[2] <- (sum(date_range$deaths))
  cumulative_cases_deaths_mortality[3] <- round(cumulative_cases_deaths_mortality[2]*100/cumulative_cases_deaths_mortality[1],2)
  
  return(cumulative_cases_deaths_mortality)
}

```

- Test function:

```{r}

#First wave
cases_between_dates("2020/05/01","2020/11/01")

#Second wave
cases_between_dates("2021/03/01","2021/09/01")
                    
#Third wave
cases_between_dates("2021/12/01","2022/01/18")

```

### Calculate confirmed cases age over time (cumulative per day & month by month)

- Create dataframe:

```{r}

#Drops unknown ages, impossible ages, unnecessary columns, and arrange by case date

avg_cases_age <- covid_CABA_confirmed %>%
  drop_na(age) %>%
  subset(age>=0 & age <=110 ) %>%
  arrange(case_date)

avg_cases_age <- avg_cases_age[,c("case_date","age")]

#Record impossible ages for accuracy

impossible_ages <- filter(covid_CABA_confirmed, age >= 110 | age < 0)

```

- Create another dataframe to calculate average age per month for confirmed cases:

```{r}

avg_age_per_month <- avg_cases_age 

#  Get months

avg_age_per_month$case_date <- as.Date(avg_age_per_month$case_date)
avg_age_per_month$month <- format(as.Date(avg_age_per_month$case_date), "%Y-%m")

#  Aggregate age on months and year and get mean
avg_age_per_month <- aggregate( age ~ month , avg_age_per_month, mean)

# Change to a nice date format (for datavis)

avg_age_per_month$month <- as.yearmon(avg_age_per_month$month)
        

```

- Calculate cumulative average age for confirmed cases day by day:

```{r}

#Calculate average per day

avg_cases_age <- avg_cases_age %>% 
 group_by(case_date) %>% 
 mutate(avg_daily_age = mean(age)) %>% 
 ungroup() %>%
 mutate(avg_cumulative = round(cumsum(avg_daily_age) / seq_along(avg_daily_age),2))

#Keep only last value for every day and add missing days

avg_cases_age <-avg_cases_age[!duplicated(avg_cases_age$case_date, fromLast = T), ]  %>%
   mutate(case_date = as.Date(case_date)) %>%
   complete(case_date = seq.Date(min(case_date), max(case_date), by="day"))

#Add missing values from last value, remove unnecessary columns

avg_cases_age$avg_cumulative <- na.locf(avg_cases_age$avg_cumulative)

avg_cases_age <- avg_cases_age[,c("case_date","avg_cumulative")]

#Match colnames to general cases table

colnames(avg_cases_age) <- c('date', 'cumulative_cases_age')

cases_by_date <- cases_by_date %>%
  left_join(avg_cases_age, by = "date")

```

### Calculate average death age over time (cumulative per day & month by month)

- Create dataframe:

```{r}

#Drops unknown ages, unnecessary columns and arrange by death date

avg_death_age <- covid_CABA_dead %>%
  drop_na(age) %>%
  arrange(death_date)

avg_death_age <- avg_death_age[,c("death_date","age")]

```

- Create another dataframe to calculate average age per month for deadly cases:

```{r}

avg_death_age_per_month <- avg_death_age 

#  Get months

avg_death_age_per_month$death_date <- as.Date(avg_death_age_per_month$death_date)
avg_death_age_per_month$month <- format(as.Date(avg_death_age_per_month$death_date), "%Y-%m")

#  Aggregate age on months and year and get mean
avg_death_age_per_month <- aggregate( age ~ month , avg_death_age_per_month, mean)

# Change to a nice date format (for datavis)

avg_death_age_per_month$month <- as.yearmon(avg_death_age_per_month$month)

```

- Join monthly tables

```{r}

# Change colnames and join monthly average age for confirmed cases and monthly average age for deadly cases tables

colnames(avg_death_age_per_month)[2] <- 'death_age'

avg_age_per_month <- avg_age_per_month %>%
  left_join(avg_death_age_per_month, by = "month")

```

- Calculate cumulative average age for deadly cases day by day:

```{r}

#Calculate mean per day

avg_death_age <- avg_death_age %>% 
 group_by(death_date) %>% 
 mutate(avg_daily_age = mean(age)) %>% 
 ungroup() %>%
 mutate(avg_cumulative = round(cumsum(avg_daily_age) / seq_along(avg_daily_age),2))

#Keep only last value for every day and add missing days 

avg_death_age <-avg_death_age[!duplicated(avg_death_age$death_date, fromLast = T), ]  %>%
   mutate(death_date = as.Date(death_date)) %>%
   complete(death_date = seq.Date(min(death_date), max(death_date), by="day"))

#Add missing values from last value, remove unnecessary columns

avg_death_age$avg_cumulative <- na.locf(avg_death_age$avg_cumulative)

avg_death_age <- avg_death_age[,c("death_date","avg_cumulative")]

#Match colnames and join with general cases table

colnames(avg_death_age) <- c('date', 'cumulative_death_age')

cases_by_date <- cases_by_date %>%
  left_join(avg_death_age, by = "date")

```


### VACCINATION

Calculate number and percentage of people vaccinated, per day, per dose

```{r}

#List amount of vaccinations by date, by number of doses and sort by date

vaccination_by_date <- vaccination_all[c(1,6,7,9)] %>%
  group_by_if(is.numeric %>% Negate) %>%
    summarize_all(sum) %>%
      arrange(date)
```

```{r}

#Complete missing days and add 0s

vaccination_by_date <- vaccination_by_date %>%
  mutate(date = as.Date(date)) %>%
  complete(date = seq.Date(min(date), max(date), by="day"))

vaccination_by_date[is.na(vaccination_by_date)] <- 0
```

```{r}
#Dose 1

vaccination_by_date <- vaccination_by_date %>%
  mutate(dose_1_cumulative = cumsum(dose_1), .after = dose_1) %>%
    mutate(dose_1_perc = round(perc(dose_1_cumulative, total_population),2), .after = dose_1_cumulative)

#Dose 2

vaccination_by_date <- vaccination_by_date %>%
  mutate(dose_2_cumulative = cumsum(dose_2), .after = dose_2) %>%
    mutate(dose_2_perc = round(perc(dose_2_cumulative, total_population),2), .after = dose_2_cumulative)

#Dose 3

vaccination_by_date <- vaccination_by_date %>%
  mutate(dose_3_cumulative = cumsum(dose_3), .after = dose_3) %>%
    mutate(dose_3_perc = round(perc(dose_3_cumulative, total_population),2), .after = dose_3_cumulative)

#Difference between 1st and 2nd dose

vaccination_by_date <- vaccination_by_date %>%
  mutate(difference_1_2 = c(vaccination_by_date$dose_1_perc - vaccination_by_date$dose_2_perc))

```


- Calculate number and vaccinations, per type, per day

```{r}

#Create a new dataframe

vaccination_by_date_by_vaccine <- vaccination_all

#Sum all vaccinations (for the same type of vaccine, gender, age, group and private/public institution) in the same day and rename col

vaccination_by_date_by_vaccine$dose_1 <- vaccination_by_date_by_vaccine$dose_1 + vaccination_by_date_by_vaccine$dose_2 + vaccination_by_date_by_vaccine$dose_3

colnames(vaccination_by_date_by_vaccine)[6]  <-  'vaccine_doses'

#Delete unnecessary cols

vaccination_by_date_by_vaccine <- vaccination_by_date_by_vaccine[c(1,4,6)] 

#Simplify vaccine names

vaccination_by_date_by_vaccine[vaccination_by_date_by_vaccine=="vacuna Coronavirus (CANSINO), vial x 1 dosis"] <- "Cansino"

vaccination_by_date_by_vaccine[vaccination_by_date_by_vaccine=="vacuna Coronavirus (PFIZER) 1ra - 2da dosis, vial x 6 dosis (0.3ml)"] <- "Pfizer"

#Sum all the vaccinations for the same date for the same type of vaccine

vaccination_by_date_by_vaccine <- vaccination_by_date_by_vaccine %>%
  group_by_if(is.numeric %>% Negate) %>%
    summarize_all(sum) %>%
      arrange(date) 

#Transpose table using vaccine names an replace NAs for days without vaccines

vaccination_by_date_by_vaccine <- vaccination_by_date_by_vaccine %>% pivot_wider(names_from = vaccine, values_from = vaccine_doses)

vaccination_by_date_by_vaccine[is.na(vaccination_by_date_by_vaccine)] <- 0

#Calculate cumulative vaccination per type of vaccine (for bar race chart)

vaccination_cumulative <- vaccination_by_date_by_vaccine

vaccination_cumulative <- as.data.frame(apply(vaccination_cumulative[,2:6],2, cumsum)) %>%
  add_column("date"=vaccination_by_date_by_vaccine$date, .before = 'Sputnik')
  

```

### Visualize some general data for CABA

```{r}

#Total population and percentage of the population infected

population_CABA = sum(population_data$population)

perc(nrow(covid_CABA_confirmed),population_CABA)

#Evolution of cases and deaths

barplot(height = cases_by_date$confirmed_cases)

barplot(height = cases_by_date$deaths)

#Total number of confirmed cases

count(covid_CABA_confirmed)

#Number of deaths

count(covid_CABA_dead)



```

### Export tables

```{r}

#Cases by date
write.csv(cases_by_date, "final_CABA_cases_by_date.csv")

#Month by month cases an death average age
write.csv(avg_age_per_month, "final_avg_age_per_month.csv")

#Vaccination by date
write.csv(vaccination_by_date,"final_vaccination_by_date.csv")

#Cumulative vaccine doses per vaccine type
write.csv(vaccination_cumulative,"final_bar_race.csv")


```

