---
title: "COVID 19 CABA // CLEANING AND PREPARING DATABASES"
output: html_notebook
---

### Load libraries

```{r}
library(tidyverse) #for all functions in tidyverse
library(lubridate) #for dates format
```

### Import all datasets

```{r}

#All Covid possible cases in CABA (Ciudad Autónoma de Buenos Aires). 
covid_all_cases <- read.csv("casos_covid19.csv", stringsAsFactors = FALSE)

#Vaccination in CABA
vaccination_all <- read.csv("dataset_total_vacunas.csv", stringsAsFactors = FALSE)

#Population and density by commune according to 2020 projection from 2010 census
population_by_commune <- read.csv("population_by_commune.csv", stringsAsFactors = FALSE)

total_population <- sum(population_by_commune$population)

#Population by neighborhood according to 2010 census
population_by_neighborhood <- read.csv("caba_pob_barrios_2010.csv", stringsAsFactors = FALSE)

#Overcrowding per commune
overcrowding_by_commune <- read.csv("overcrowding.csv", stringsAsFactors = FALSE)

#Average income for +10 years old population
income_plus_10_by_commune <- read.csv("income_plus_10.csv", stringsAsFactors = FALSE)

#Poverty percentage per commune
poverty_by_commune <- read.csv("poverty.csv", stringsAsFactors = FALSE)

#Neighborhood per commune
neighborhood_by_commune <- read.csv("barrios_comunas.csv", stringsAsFactors = FALSE)


```


### Cleaning and preparing datasets

- Check datasets for Covid cases and vaccination:

```{r}
head(covid_all_cases) # Column names are in Spanish

head(vaccination_all) # Column names are already translated to English 
```

- Keep only necessary columns, and change name to English:

```{r}
covid_all_cases <- covid_all_cases[c(2,5,6,7,8,9,10,11,12)]

colnames(covid_all_cases) <-c("case_date","region","neighborhood", "commune","gender","age","case_type","death_date","dead")

colnames(population_by_neighborhood) <-c("neighborhood", "population")

colnames(neighborhood_by_commune) <-c("neighborhood", "commune")

```

- Fix dates format for cases, deaths and vaccination:

```{r}

#For Covid cases and deaths

covid_all_cases$case_date <- substr(covid_all_cases$case_date,1,nchar(covid_all_cases$case_date)-16)

covid_all_cases$death_date <- substr(covid_all_cases$death_date,1,nchar(covid_all_cases$death_date)-16)

covid_all_cases$case_date <- dmy(covid_all_cases$case_date)

covid_all_cases$death_date <- dmy(covid_all_cases$death_date)

#For vaccination

vaccination_all$date <- substr(vaccination_all$date,1,nchar(vaccination_all$date)-9)

vaccination_all$date <- dmy(vaccination_all$date)

```


- Create a dataframe with possible cases from CABA residents:

```{r}
covid_CABA_cases <- subset(covid_all_cases, covid_all_cases$region == 'CABA')

```

- Create a dataframe for confirmed cases and another subset for deadly confirmed cases:

*I subset the latter from covid_CABA_confirmed and not from covid_CABA_cases because there are people who were tested and died with suspected Covid, but it was later confirmed that their death was due to other causes.*

```{r}

covid_CABA_confirmed <- subset(covid_CABA_cases, covid_CABA_cases$case_type == 'confirmado')

covid_CABA_dead <-subset(covid_CABA_confirmed, covid_CABA_confirmed$dead == 'si')
```

- Create a dataframe with socio-economic and demographic data by commune

```{r}

population_data <- population_by_commune %>%
  left_join(overcrowding_by_commune, by = "commune") %>%
  left_join(poverty_by_commune, by = "commune") %>%
  left_join(income_plus_10_by_commune, by = "commune")

```

### Export dataframes as CSVs

```{r}
write.csv(covid_CABA_dead, "clean_covid_CABA_dead.csv")
write.csv(covid_CABA_confirmed, "clean_covid_CABA_confirmed.csv")
write.csv(population_data, "clean_population_data.csv")
write.csv(vaccination_all, "clean_vaccination.csv")
```
